---
title: "Reproducable Research - Assignment 1"
author: "Arjit"
date: "October 13, 2015"
output: html_document
---

Setting echo = true
```{r setoptions,echo=TRUE,results="hold"}

```

Downloading the Dataset and Formatting the date value from character to date
and converting interval field to factor

```{r}
library(ggplot2)
act_data<-read.csv(file = "C:/Users/Nemes!s/Documents/GitHub/RepData_PeerAssessment1/activity.csv ",sep=",",header=TRUE,colClasses = c("numeric","character","numeric"))
str(act_data)
act_data$date<- as.Date(act_data$date,"%Y-%m-%d")
act_data$interval <- as.factor(act_data$interval)
str(act_data)

```
Creating a new Dataframe to aggregate day wise steps data

```{r}
new_dat <- aggregate(steps ~ date, act_data, sum)
colnames(new_dat)<-c("date","steps")
```

Plotting the historogram for no of steps taken per day

```{r}
hist(new_dat$steps,breaks=10,col="blue",main="No of Steps each day",xlab = "Steps")
```

calculating the mean and median of steps

```{r}
dat_mean<-mean(new_dat$steps,na.rm=TRUE)
dat_median<- median(new_dat$steps,na.rm=TRUE)
```
The mean is `r dat_mean` and median is `rdat_median`.

Calculating  the average daily pattern

````{r}
av_dat<-aggregate(act_data$steps, by= list(interval=act_data$interval), FUN=mean,na.rm=TRUE)

av_dat$interval <- 
        as.integer(levels(av_dat$interval)[av_dat$interval])
colnames(av_dat)<-c("interval","steps")
ggplot(av_dat,aes(x=interval , y= steps))+geom_line(color="green") + labs( title="Average Daily Pattern", x="Interval", y="No of Steps")

````
Finding  the 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
`````{r}
max_interval<-av_dat[which.max(av_dat$steps), ] 

`````

The max inerval for the highest step takes place at `max_interval`

Calculating the total number of missing values

`````{r}
no_NA_nos<- sum(is.na(act_data$steps))
          
``````
Total number of NAs is `r no_NA_nos`

Replacing NAs with mean of steps with mean of that day

````{r}
na_fill <- function(act_data, pervalue) {
        na_index <- which(is.na(act_data$steps))
        na_replace <- unlist(lapply(na_index, FUN=function(idx){
                interval = act_data[idx,]$interval
                pervalue[pervalue$interval == interval,]$steps
        }))
        fill_steps <- act_data$steps
        fill_steps[na_index] <- na_replace
        fill_steps
}

act_data_fill <- data.frame(  
        steps = na_fill(act_data, av_dat),  
        date = act_data$date,  
        interval = act_data$interval)
str(act_data_fill)

`````
Plotting the graph

````{r}
fill_steps_per_day <- aggregate(steps ~ date, act_data_fill, sum)
colnames(fill_steps_per_day) <- c("date","steps")

##plotting the histogram
ggplot(fill_steps_per_day, aes(x = steps)) + 
       geom_histogram(fill = "green", binwidth = 1000) + 
        labs(title="Histogram of Steps Taken per Day", 
             x = "Number of Steps per Day", y = "Number of times in a day(Count)")


``````
Calculating and reporting the mean and median

`````{r}

mean_2  <- mean(fill_steps_per_day$steps, na.rm=TRUE)
median_2 <- median(fill_steps_per_day$steps, na.rm=TRUE)

```
The mean value is `r mean_2` and median is `r median_2`




Differences in activity patterns between weekdays and weekends

Step1: Adding a column to indicate day of the week 

````{r}
weekdays_steps <- function(data) {
    weekdays_steps <- aggregate(data$steps, by=list(interval = data$interval),
                          FUN=mean, na.rm=T)
    # convert to integers for plotting
    weekdays_steps$interval <- 
            as.integer(levels(weekdays_steps$interval)[weekdays_steps$interval])
    colnames(weekdays_steps) <- c("interval", "steps")
    weekdays_steps
}

data_by_weekdays <- function(data) {
    data$weekday <- 
            as.factor(weekdays(data$date)) # weekdays
    weekend_data <- subset(data, weekday %in% c("Saturday","Sunday"))
    weekday_data <- subset(data, !weekday %in% c("Saturday","Sunday"))

    weekend_steps <- weekdays_steps(weekend_data)
    weekday_steps <- weekdays_steps(weekday_data)

    weekend_steps$dayofweek <- rep("weekend", nrow(weekend_steps))
    weekday_steps$dayofweek <- rep("weekday", nrow(weekday_steps))

    data_by_weekdays <- rbind(weekend_steps, weekday_steps)
    data_by_weekdays$dayofweek <- as.factor(data_by_weekdays$dayofweek)
    data_by_weekdays
}

data_weekdays <- data_by_weekdays(act_data_fill)
````

Plotting the graph

````{r}
ggplot(data_weekdays, aes(x=interval, y=steps)) + 
        geom_line(color="violet") + 
        facet_wrap(~ dayofweek, nrow=2, ncol=1) +
        labs(x="Interval", y="Number of steps") +
        theme_bw()

```